Report on Certificate Tools

### 2016-11-16 TEX/draft.tex
Started Conclusion, ommited Notes -> Do a short Review and their + and - in a Conclusion part

\section{General}
%TABLE START
\begin{table}[h!]
\centering
\caption{Overwiew General}
\label{my-label}
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{|c|c|c|c|c|}
\hline
                        & \multicolumn{4}{c|}{General}                                                                               \\ \cline{2-5} 
\multirow{-2}{*}{Tools} & Type   & View + info                                        & License & Operating System                   \\ \hline
OpenSSL                 & PKCS12 & \cellcolor[HTML]{34FF34}Yes                        & Public  & \cellcolor[HTML]{34FF34}Both       \\ \hline
Keytool                 & JKS    & \cellcolor[HTML]{34FF34}{\color[HTML]{000000} Yes} & Public  & \cellcolor[HTML]{34FF34}Both       \\ \hline
GnuPG                   &        &                                                    &         &                                    \\ \hline
Keystore Explorer       & Both   & \cellcolor[HTML]{34FF34}Yes                        & Public  & \cellcolor[HTML]{FFC702}Win | OS X \\ \hline
XCA                     & PKCS12 & \cellcolor[HTML]{34FF34}Yes                        & Public  & \cellcolor[HTML]{34FF34}Both       \\ \hline
GnuTLS                  & PKCS12 & \cellcolor[HTML]{34FF34}Yes                        & Public  & \cellcolor[HTML]{FFC702}LINUX      \\ \hline
Makecert                & PKCS12 & \cellcolor[HTML]{FE0000}No			      & Public  & \cellcolor[HTML]{FFC702}Windows    \\ \hline
\end{tabular}
\end{adjustbox}
\end{table}
%TABLE END

@@

  \subsection{Type}
  OpenSSL uses PKCS12 to store keys and or certificates. Certificates and keys  made in OpenSSL however are being made into PEM or DER encoded file. Said keys/certificates can then be stored inside single .pfx file. 
  \cellcolor[HTML]{34FF34}Yes             
  \subsection{View and information}

@@

      \paragraph{Self-signed certificate}
      It is possible to generate private key and self signed certificate with 
      \begin{verbatim}
      openssl req -x509 -sha256 -nodes -days 365 \
	      -newkey rsa:[length] -keyout privateKey.key \
              -out certificate_name.crt
      \end{verbatim}
      \cellcolor[HTML]{34FF34}Yes             
      \paragraph{Certificate Signing Request}
      It is possible to generate private key and CSR with
      \begin{verbatim}
      openssl req -out CSR.csr -new -newkey rsa:[length] \
              -nodes -keyout privateKey.key
      \end{verbatim}

@@

@@

    \subsubsection{Support for Cryptographic Service Provider}
Support for Cryptographic Service Provider is native but there is need for a definition of path where the CSP is located.
\begin{verbatim}
-storetype [keystore_type_of_implementation]
-provider [provider_implementation_name]
-providerpath [provider_implementation_type]
\end{verbatim}
%unsure about that but http://stackoverflow.com/questions/6866447/keytool-error-java-security-keystoreexception-bks-not-found

@@

  \subsection{Type}
The type of file with which GnuPG works is a .gpg encoded in OpenPGP standard. As I understood the OpenPGP is competing standard to X509 standard.  
  \subsection{View and information}

@@

      \paragraph{Self-signed certificate}
Generation of keys and self-signed certificate is not supported. In older versions it was possigle with \verb+gpgm+
      \paragraph{Certificate Signing Request}

@@

More options when generating self-signed certificate (or a CSR) you need to create a config file which has structure as follows (copied from a man page):
\begin{verbatim}
# X.509 Certificate options
#
# DN options

# The organization of the subject.
organization = "Koko inc."

# The organizational unit of the subject.
unit = "sleeping dept."

# The locality of the subject.
# locality =

# The state of the certificate owner.
state = "Attiki"

# The country of the subject. Two letter code.
country = GR

# The common name of the certificate owner.
cn = "Cindy Lauper"

# A user id of the certificate owner.
#uid = "clauper"

# Set domain components
#dc = "name"
#dc = "domain"

# If the supported DN OIDs are not adequate you can set
# any OID here.
# For example set the X.520 Title and the X.520 Pseudonym
# by using OID and string pairs.
#dn_oid = "2.5.4.12 Dr."
#dn_oid = "2.5.4.65 jackal"

# This is deprecated and should not be used in new
# certificates.
# pkcs9_email = "none@none.org"

# An alternative way to set the certificate's distinguished name directly
# is with the "dn" option. The attribute names allowed are:
# C (country), street, O (organization), OU (unit), title, CN (common name),
# L (locality), ST (state), placeOfBirth, gender, countryOfCitizenship, 
# countryOfResidence, serialNumber, telephoneNumber, surName, initials, 
# generationQualifier, givenName, pseudonym, dnQualifier, postalCode, name, 
# businessCategory, DC, UID, jurisdictionOfIncorporationLocalityName, 
# jurisdictionOfIncorporationStateOrProvinceName,
# jurisdictionOfIncorporationCountryName, XmppAddr, and numeric OIDs.

#dn = "cn = Nikos,st = New\, Something,C=GR,surName=Mavrogiannopoulos,2.5.4.9=Arkadias"

# The serial number of the certificate
# Comment the field for a time-based serial number.
serial = 007

# In how many days, counting from today, this certificate will expire.
# Use -1 if there is no expiration date.
expiration_days = 700

# Alternatively you may set concrete dates and time. The GNU date string 
# formats are accepted. See:
# http://www.gnu.org/software/tar/manual/html_node/Date-input-formats.html

#activation_date = "2004-02-29 16:21:42"
#expiration_date = "2025-02-29 16:24:41"

# X.509 v3 extensions

# A dnsname in case of a WWW server.
#dns_name = "www.none.org"
#dns_name = "www.morethanone.org"

# An othername defined by an OID and a hex encoded string
#other_name = "1.3.6.1.5.2.2 302ca00d1b0b56414e5245494e2e4f5247a11b3019a006020400000002a10f300d1b047269636b1b0561646d696e"
#other_name_utf8 = "1.2.4.5.6 A UTF8 string"
#other_name_octet = "1.2.4.5.6 A string that will be encoded as ASN.1 octet string"

# Allows writing an XmppAddr Identifier
#xmpp_name = juliet@im.example.com

# Names used in PKINIT
#krb5_principal = user@REALM.COM
#krb5_principal = HTTP/user@REALM.COM

# A subject alternative name URI
#uri = "http://www.example.com"

# An IP address in case of a server.
#ip_address = "192.168.1.1"

# An email in case of a person
email = "none@none.org"

# TLS feature (rfc7633) extension. That can is used to indicate mandatory TLS
# extension features to be provided by the server. In practice this is used
# to require the Status Request (extid: 5) extension from the server. That is,
# to require the server holding this certificate to provide a stapled OCSP response.
# You can have multiple lines for multiple TLS features.

# To ask for OCSP status request use:
#tls_feature = 5

@@

# Password when encrypting a private key
#password = secret

# An URL that has CRLs (certificate revocation lists)
# available. Needed in CA certificates.
#crl_dist_points = "http://www.getcrl.crl/getcrl/"

# Whether this is a CA certificate or not
#ca

# Subject Unique ID (in hex)
#subject_unique_id = 00153224

# Issuer Unique ID (in hex)
#issuer_unique_id = 00153225

#### Key usage

# The following key usage flags are used by CAs and end certificates

# Whether this certificate will be used to sign data (needed
# in TLS DHE ciphersuites). This is the digitalSignature flag
# in RFC5280 terminology.
signing_key

@@

# Whether this key will be used to sign other certificates. The
# keyCertSign flag in RFC5280 terminology.
#cert_signing_key

# Whether this key will be used to sign CRLs. The
# cRLSign flag in RFC5280 terminology.
#crl_signing_key

# The keyAgreement flag of RFC5280. It's purpose is loosely
# defined. Not use it unless required by a protocol.
#key_agreement

# The dataEncipherment flag of RFC5280. It's purpose is loosely
# defined. Not use it unless required by a protocol.
#data_encipherment

# The nonRepudiation flag of RFC5280. It's purpose is loosely
# defined. Not use it unless required by a protocol.
#non_repudiation

#### Extended key usage (key purposes)

# The following extensions are used in an end certificate
# to clarify its purpose. Some CAs also use it to indicate
# the types of certificates they are purposed to sign.

# Whether this certificate will be used for a TLS client;
# this sets the id-kp-serverAuth (1.3.6.1.5.5.7.3.1) of 
# extended key usage.
#tls_www_client

# Whether this certificate will be used for a TLS server;
# This sets the id-kp-clientAuth (1.3.6.1.5.5.7.3.2) of 
# extended key usage.
#tls_www_server

# Whether this key will be used to sign code. This sets the
# id-kp-codeSigning (1.3.6.1.5.5.7.3.3) of extended key usage
# extension.
#code_signing_key

# Whether this key will be used to sign OCSP data. This sets the
# id-kp-OCSPSigning (1.3.6.1.5.5.7.3.9) of extended key usage extension.
#ocsp_signing_key

# Whether this key will be used for time stamping. This sets the
# id-kp-timeStamping (1.3.6.1.5.5.7.3.8) of extended key usage extension.
#time_stamping_key

# Whether this key will be used for email protection. This sets the
# id-kp-emailProtection (1.3.6.1.5.5.7.3.4) of extended key usage extension.
#email_protection_key

# Whether this key will be used for IPsec IKE operations (1.3.6.1.5.5.7.3.17).
#ipsec_ike_key

## adding custom key purpose OIDs

# for microsoft smart card logon
# key_purpose_oid = 1.3.6.1.4.1.311.20.2.2

# for email protection
# key_purpose_oid = 1.3.6.1.5.5.7.3.4

# for any purpose (must not be used in intermediate CA certificates)
# key_purpose_oid = 2.5.29.37.0

### end of key purpose OIDs

### Adding arbitrary extensions
# This requires to provide the extension OIDs, as well as the extension data in
# hex format. The following two options are available since GnuTLS 3.5.3.
#add_extension = "1.2.3.4 0x0AAB01ACFE"

# As above but encode the data as an octet string
#add_extension = "1.2.3.4 octet_string(0x0AAB01ACFE)"

# For portability critical extensions shouldn't be set to certificates.
#add_critical_extension = "5.6.7.8 0x1AAB01ACFE"

# When generating a certificate from a certificate
# request, then honor the extensions stored in the request
# and store them in the real certificate.
#honor_crq_extensions

# Alternatively only specific extensions can be copied.
#honor_crq_ext = 2.5.29.17
#honor_crq_ext = 2.5.29.15

# Path length contraint. Sets the maximum number of
# certificates that can be used to certify this certificate.
# (i.e. the certificate chain length)
#path_len = -1
#path_len = 2

# OCSP URI
# ocsp_uri = http://my.ocsp.server/ocsp

# CA issuers URI
# ca_issuers_uri = http://my.ca.issuer

# Certificate policies
#policy1 = 1.3.6.1.4.1.5484.1.10.99.1.0
#policy1_txt = "This is a long policy to summarize"
#policy1_url = http://www.example.com/a-policy-to-read

#policy2 = 1.3.6.1.4.1.5484.1.10.99.1.1
#policy2_txt = "This is a short policy"
#policy2_url = http://www.example.com/another-policy-to-read

# Name constraints

# DNS
#nc_permit_dns = example.com
#nc_exclude_dns = test.example.com

# EMAIL
#nc_permit_email = "nmav@ex.net"

# Exclude subdomains of example.com
#nc_exclude_email = .example.com

# Exclude all e-mail addresses of example.com
#nc_exclude_email = example.com

# IP
#nc_permit_ip = 192.168.0.0/16
#nc_exclude_ip = 192.168.5.0/24
#nc_permit_ip = fc0a:eef2:e7e7:a56e::/64

# Options for proxy certificates
#proxy_policy_language = 1.3.6.1.5.5.7.21.1

# Options for generating a CRL

# The number of days the next CRL update will be due.
# next CRL update will be in 43 days
#crl_next_update = 43

# this is the 5th CRL by this CA
# Comment the field for a time-based number.
#crl_number = 5

# Specify the update dates more precisely.
#crl_this_update_date = "2004-02-29 16:21:42"
#crl_next_update_date = "2025-02-29 16:24:41"

# The date that the certificates will be made seen as
# being revoked.
#crl_revocation_date = "2025-02-29 16:24:41"
\end{verbatim}

@@

>>>>

\section{General}
\begin{table}[h!]
\centering
\caption{Overwiew General}
\label{my-label}
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{|c|c|c|c|c|}
\hline
                        & \multicolumn{4}{c|}{General}                                                                               \\ \cline{2-5} 
\multirow{-2}{*}{Tools} & Type   & View + info                                        & License & Operating System                   \\ \hline
OpenSSL                 & PKCS12 & \cellcolor[HTML]{34FF34}Yes                        & Public  & \cellcolor[HTML]{34FF34}Both       \\ \hline
Keytool                 & JKS    & \cellcolor[HTML]{34FF34}{\color[HTML]{000000} Yes} & Public  & \cellcolor[HTML]{34FF34}Both       \\ \hline
GnuPG                   &        &                                                    &         &                                    \\ \hline
Keystore Explorer       & Both   & \cellcolor[HTML]{34FF34}Yes                        & Public  & \cellcolor[HTML]{FFC702}Win &\vert& OS X \\ \hline
XCA                     & PKCS12 & \cellcolor[HTML]{34FF34}Yes                        & Public  & \cellcolor[HTML]{34FF34}Both       \\ \hline
GnuTLS                  & PKCS12 & \cellcolor[HTML]{34FF34}Yes                        & Public  & \cellcolor[HTML]{FFC702}LINUX      \\ \hline
Makecert                & PKCS12 & \cellcolor[HTML]{FE0000}No			      & Public  & \cellcolor[HTML]{FFC702}Windows    \\ \hline
\end{tabular}
\end{adjustbox}
\end{table}
%TABLE END

@@

  \subsection{Type}
  OpenSSL uses PKCS12 to store keys and or certificates. Certificates and keys  made in OpenSSL however are being made into PEM or DER encoded file. Said keys/certificates can then be stored inside single .pfx file.

@@

      \paragraph{Self-signed certificate}
      It is possible to generate private key and self signed certificate with 
      \begin{verbatim}
      openssl req -x509 -sha256 -nodes -days 365 \
	      -newkey rsa:[length] -keyout privateKey.key \
              -out certificate_name.crt
      \end{verbatim}

      \paragraph{Certificate Signing Request}
      It is possible to generate private key and CSR with
      \begin{verbatim}
      openssl req -out CSR.csr -new -newkey rsa:[length] \
              -nodes -keyout privateKey.key
      \end{verbatim}

@@

\section{Notes}

@@

    \subsubsection{Support for Cryptographic Service Provider}
Support for Cryptographic Service Provider is native but there is need for a definition of path where the CSP is located.
\begin{verbatim}
-storetype [keystore_type_of_implementation]
-provider [provider_implementation_name]
-providerpath [provider_implementation_type]
\end{verbatim}

@@

  \subsection{Type}
GnuPG can make and view X509 Certificate Signing Requests
  \subsection{View and information}

@@

      \paragraph{Self-signed certificate}
Generation of keys and self-signed certificate is not supported. In older versions it was possigle with \verb+gpgm+
      \paragraph{Certificate Signing Request}
Possible, but private key is not a part of output CSR.      
      \paragraph{Specify length}

@@

More options when generating self-signed certificate (or a CSR) you need to create a config file which has structure as follows (copied from a man page):
\begin{verbatim}

dn = "cn=Common Name,st=State Name,C=CR,O=Organization,OU=OrganizationUnit,L=Locality"
serial = <serial_nr>
expiration_days = <int_days_from_now>

# X.509 v3 extensions
dns_name = "www.dns1.org"
dns_name = "www.dns2.org"
...

ip_address = "192.168.1.1"

@@

# Password when encrypting a private key
password = secret

# Whether this is a CA certificate or not
ca

@@

# Whether this key will be used to sign other certificates. The
# keyCertSign flag in RFC5280 terminology.
cert_signing_key

# When generating a certificate from a certificate
# request, then honor the extensions stored in the request
# and store them in the real certificate.
honor_crq_extensions

# Path length contraint. Sets the maximum number of
# certificates that can be used to certify this certificate.
# (i.e. the certificate chain length)
#path_len = -1
#path_len = 2
\end{verbatim}

@@

\part{Conclusion}

\chapter{Tools Review}

The main thing to understand is, that every tool has had its strong and weak points. Every should be used in complience with required properties of use. That can mean restrictions on used Operating System (e.g. customer has only access to Apple computers with OS X operating system), or for example restrictions how it should be used on Operating System (command line vs GUI programm). In the next part i will try to go over strengths and weaknesses of each tool that I have tested.

Section{}
